---
title: "Section_02 Exercises"
output: html_notebook
---

A collaborator sends you a file containing data for two years of average race finish times, "times.csv":

age_group,2015_time,2015_participants,2016_time,2016_participants
20,3:46,54,3:22,62
30,3:50,60,3:43,58
40,4:39,29,3:49,33
50,4:48,10,4:59,14
You read in the data file:

d <- read_csv("times.csv")
```{r}
library (tidyverse)
```

```{r}
d <- read_csv("C:/Users/lfima/Documents/edx/06_Data Wrangling/Section_02_Reshape/data/times.csv")
```
Tidying data:
```{r} 
d %>% gather(key = "key", value = "value", -age_group) %>% separate(key, into = c("year", "variable_name"), sep = "_") %>% spread(variable_name, value = value)
```

---


You are in the process of tidying some data on heights, hand length, and wingspan for basketball players in the draft. Currently, you have the following:

    > head(stats)
key               value
allen_height      75
allen_hand_length 8.25
allen_wingspan	  79.25
bamba_height      83.25
bamba_hand_length 9.75
bamba_wingspan    94
  
Select all of the correct commands below that would turn this data into a “tidy” format with columns "height", "hand_length" and "wingspan".
```{r}
n <- read_csv("C:/Users/lfima/Documents/edx/06_Data Wrangling/Section_02_Reshape/data/nba.csv")
n
```

```{r}
n %>% separate(col = key, into = c("name", "variables"), sep = "_")
```
```{r}
n %>% separate(col = key, into = c("name", "variables"), sep = "_", extra = "merge") %>% spread(key = variables, value = value)
```

## Assessment 02

Use the following libraries for these questions:


```{r}
library(tidyverse)
library(dslabs)
library (ggplot2)
```


### Question 9
1 point possible (graded)
Examine the built-in dataset co2. This dataset comes with base R, not dslabs - just type co2 to access the dataset.

```{r}
co2
```

### Question 10
1 point possible (graded)
Run the following command to define the co2_wide object:

```{r}
co2_wide <- data.frame(matrix(co2, ncol = 12, byrow = TRUE)) %>% 
      setNames(1:12) %>%
    mutate(year = as.character(1959:1997))
head(co2_wide)
```

Use the gather() function to make this dataset tidy. Call the column with the CO2 measurements co2 and call the month column month. Name the resulting object co2_tidy.

Which code would return the correct tidy format?
```{r}
co2_tidy <- co2_wide %>% gather(key = "month",value = "co2","1":"12")
co2_tidy
```

Question 11
1 point possible (graded)
Use co2_tidy to plot CO2 versus month with a different curve for each year:

```{r}
co2_tidy %>% ggplot(aes(as.numeric(month), co2, color = year)) + geom_line()

```

What can be concluded from this plot? <br>
CO2 concentrations are highest around May and the yearly average increased from 1959 to 1997.<br>

---

### Question 12
Load the admissions dataset from dslabs, which contains college admission information for men and women across six majors, and remove the applicants percentage column:

```{r}
library(dslabs)
data(admissions)
dat <- admissions %>% select(-applicants)
head(dat)
```

Your goal is to get the data in the shape that has one row for each major, like this:
```{}
major  men   women
A      62    82		
B      63    68		
C      37    34		
D      33    35		
E      28    24		
F       6     7	
```
Which command could help you to wrangle the data into the desired format?
```{r}
dat %>% spread (gender, value = admitted)
```

---

### Question 13
Now use the admissions dataset to create the object tmp, which has columns major, gender, key and value:

```{r}
tmp <- gather(admissions, key, value, admitted:applicants)
tmp
```
    

Combine the key and gender and create a new column called column_name to get a variable with the following values: admitted_men, admitted_women, applicants_men and applicants_women. Save the new data as tmp2.

Which command could help you to wrangle the data into the desired format?
```{r}
tmp2 <- tmp %>% unite(column_name, key, gender)
tmp2
```

---

### Question 14
1 point possible (graded)
Which function can reshape tmp2 to a table with six rows and five columns named major, admitted_men, admitted_women, applicants_men and applicants_women?

```{}
major   admitted_men    admitted_women    applicants_men    applicants_women
A                              
B
```

```{r}
tmp2 %>% spread(key = column_name, value = value)
```

---

## Section 2.2
```{r}
library (dplyr)
```

## Question 1
1 point possible (graded)
You have created data frames tab1 and tab2 of state population and election data, similar to our module videos:

```{r}
tab1 <- data.frame(state = c("Alabama", "Alaska", "Arizona", "Delaware","District of Columbia"), population = c(4779736,710231,6392017, 897934, 601723  ))
tab1
```

```{r}
tab2 <- data.frame(state = c("Alabama", "Alaska", "Arizona", "California","Colorado", "Connecticut"), population = c(9,3,11, 55, 9,7  ))
tab2
```
What are the dimensions of the table dat, created by the following command?
dat <- left_join(tab1, tab2, by = “state”)
```{r}
dim(left_join(tab1, tab2, by = "state"))
```
Question 2
1 point possible (graded)
We are still using the tab1 and tab2 tables shown in question 1. What join command would create a new table “dat” with three rows and two columns?
      
```{r}
dim(semi_join(tab1, tab2, by = "state") )
```


## Introduction to Questions 5-7
Install and load the Lahman library. This library contains a variety of datasets related to US professional baseball. We will use this library for the next few questions and will discuss it more extensively in the Regression course. For now, focus on wrangling the data rather than understanding the statistics.

The Batting data frame contains the offensive statistics for all baseball players over several seasons.  Filter this data frame to define top as the top 10 home run (HR) hitters in 2016:
```{r}
install.packages("Lahman")
```
```{r}
package_path <- system.file("extdata", package="Lahman")
list.files(package_path)
```


```{r}
library(Lahman)
library (tidyverse)

top <- Batting %>% 
  filter(yearID == 2016) %>%
  arrange(desc(HR)) %>%    # arrange by descending HR count
  slice(1:10)    # take entries 1-10
top %>% as_tibble()
```


Also Inspect the Master data frame, which has demographic information for all players:
```{r}
People %>% as_tibble()
```

## Question 5
1 point possible (graded)
Use the correct join or bind function to create a combined table of the names and statistics of the top 10 home run (HR) hitters for 2016. This table should have the player ID, first name, last name, and number of HR for the top 10 players. Name this data frame top_names.

Identify the join or bind that fills the blank in this code to create the correct table:


top_names <- top %>% ___________________ %>%
    select(playerID, nameFirst, nameLast, HR)


Which bind or join function fills the blank to generate the correct table?
```{r}
top_names <- top %>% left_join(People) %>% select(playerID, nameFirst, nameLast, HR)
top_names
```

## Question 6
1 point possible (graded)
Inspect the Salaries data frame. Filter this data frame to the 2016 salaries, then use the correct bind join function to add a salary column to the top_names data frame from the previous question. Name the new data frame top_salary. Use this code framework:

```{r}
Salaries %>% filter(yearID == 2016) 
```


top_salary <- Salaries %>% filter(yearID == 2016) %>%
  ______________ %>%
  select(nameFirst, nameLast, teamID, HR, salary)

    
Which bind or join function fills the blank to generate the correct table?

```{r}
top_salary <- Salaries %>% filter(yearID == 2016) %>% right_join(top_names) %>%  select(nameFirst, nameLast, teamID, HR, salary)
top_salary
```


## Question 7
2 points possible (graded)
Inspect the AwardsPlayers table. Filter awards to include only the year 2016.
```{r}
awards_2016 <- AwardsPlayers %>% filter(yearID == 2016) %>% select(playerID)
awards_2016

top_names_2016 <- top_names %>% select(playerID)
top_names_2016

```

How many players from the top 10 home run hitters won at least one award in 2016? - Use set operators
```{r}
intersect(awards_2016, top_names_2016)
```



```{r}
inner_join(top_names, awards_2016, by = "playerID")
```

How many players won an award in 2016 but were not one of the top 10 home run hitters in 2016?
Use a set operator.
```{r}
setdiff(awards_2016, top_names_2016)
```

## Assessment: Web Scraping

### Question 01

```{r}
library(rvest)
url <- "https://web.archive.org/web/20181024132313/http://www.stevetheump.com/Payrolls.htm"
h <- read_html(url)
```

We learned that tables in html are associated with the table node.  Use the html_nodes() function and the table node type to extract the first table. Store it in an object nodes:
```{r}
nodes <- html_nodes(h, "table")
nodes
```


The html_nodes() function returns a list of objects of class xml_node. We can see the content of each one using, for example, the html_text() function. You can see the content for an arbitrarily picked component like this:

```{r}
class (nodes[[8]])
```


If the content of this object is an html table, we can use the html_table() function to convert it to a data frame:
```{r}
html_table(nodes[[8]])
```


You will analyze the tables from this HTML page over questions 1-3.

## Question 1
0.0/2.5 points (graded)
Many tables on this page are team payroll or salary information tables - with columns for rank, team, and one or more money values.

Convert the first four tables in nodes to data frames and inspect them.

Note that "parsing errors" and/or "empty tables" still count towards the table index!

Which of the first four nodes are tables of team payroll?
Check all correct answers. Look at table content, not column names.

Remember that payroll is information about the total salary on a team.
```{r}
#nodes <- html_nodes(h, "table")
#html_table(nodes[[1]])
html_table(nodes[[2]])
#html_table(nodes[[3]])
#html_table(nodes[[4]])
```



## Question 2
0.0/2.0 points (graded)
For the last 3 components of nodes, which of the following are true? (Check all correct answers.)
Check all correct answers.


```{r}
#html_table(nodes[[19]])
#html_table(nodes[[20]])
html_table(nodes[[21]])

```


## Question 3
1 point possible (graded)
Create a table called tab_1 using entry 10 of nodes. Create a table called tab_2 using entry 19 of nodes.

```{r}
tab1 <- html_table(nodes[[10]])
tab1
tab2 <- html_table(nodes[[19]])
tab2
```
Note that the column names should be c("Team", "Payroll", "Average"). You can see that these column names are actually in the first data row of each table, and that tab_1 has an extra first column No. that should be removed so that the column names for both tables match.

Remove the extra column in tab_1, remove the first row of each dataset, and change the column names for each table to c("Team", "Payroll", "Average"). Use a full_join() by the Team to combine these two tables.
```{r}
tab1 <- slice(tab1, 2:nrow(tab1))
tab1 <- tab1[,c(2:4)]
tab1 <- set_names(tab1, c("Team", "Payroll", "Average"))
tab1
```
```{r}
tab2 <- slice(tab2, 2:nrow(tab2))
tab2 <- set_names(tab2, c("Team", "Payroll", "Average"))
tab2
```
```{r}
full_join(tab1,tab2, by="Team")
```
How many rows are in the joined data table?
```{r}
nrow(full_join(tab1,tab2, by="Team"))
```

## Introduction: Questions 4 and 5
The Wikipedia page on opinion polling for the Brexit referendum External link, in which the United Kingdom voted to leave the European Union in June 2016, contains several tables. One table contains the results of all polls regarding the referendum over 2016:

```{r}
library(rvest)
library(tidyverse)
url <- "https://en.wikipedia.org/w/index.php?title=Opinion_polling_for_the_United_Kingdom_European_Union_membership_referendum&oldid=896735054"
```

## Question 4

Assign tab to be the html nodes of the "table" class.

How many tables are in this Wikipedia page?
```{r}
p <- read_html(url)
nodes <- html_nodes(p, "table")
nodes
```


## Question 5
1 point possible (graded)
Inspect the first several html tables using html_table() with the argument fill=TRUE (you can read about this argument in the documentation). Find the first table that has 9 columns with the first column named "Date(s) conducted".
```{r}
ncol(html_table(nodes[[6]], fill=TRUE))
```
```{r}
html_table(nodes[[6]])
```

What is the first table number to have 9 columns where the first column is named "Date(s) conducted"?
```{r}
6
```

