---
title: "Section_02 Exercises"
output: html_notebook
---

A collaborator sends you a file containing data for two years of average race finish times, "times.csv":

age_group,2015_time,2015_participants,2016_time,2016_participants
20,3:46,54,3:22,62
30,3:50,60,3:43,58
40,4:39,29,3:49,33
50,4:48,10,4:59,14
You read in the data file:

d <- read_csv("times.csv")
```{r}
library (tidyverse)
```

```{r}
d <- read_csv("C:/Users/lfima/Documents/edx/06_Data Wrangling/Section_02_Reshape/data/times.csv")
```
Tidying data:
```{r} 
d %>% gather(key = "key", value = "value", -age_group) %>% separate(key, into = c("year", "variable_name"), sep = "_") %>% spread(variable_name, value = value)
```

---


You are in the process of tidying some data on heights, hand length, and wingspan for basketball players in the draft. Currently, you have the following:

    > head(stats)
key               value
allen_height      75
allen_hand_length 8.25
allen_wingspan	  79.25
bamba_height      83.25
bamba_hand_length 9.75
bamba_wingspan    94
  
Select all of the correct commands below that would turn this data into a “tidy” format with columns "height", "hand_length" and "wingspan".
```{r}
n <- read_csv("C:/Users/lfima/Documents/edx/06_Data Wrangling/Section_02_Reshape/data/nba.csv")
n
```

```{r}
n %>% separate(col = key, into = c("name", "variables"), sep = "_")
```
```{r}
n %>% separate(col = key, into = c("name", "variables"), sep = "_", extra = "merge") %>% spread(key = variables, value = value)
```

## Assessment 02

Use the following libraries for these questions:


```{r}
library(tidyverse)
library(dslabs)
library (ggplot2)
```


### Question 9
1 point possible (graded)
Examine the built-in dataset co2. This dataset comes with base R, not dslabs - just type co2 to access the dataset.

```{r}
co2
```

### Question 10
1 point possible (graded)
Run the following command to define the co2_wide object:

```{r}
co2_wide <- data.frame(matrix(co2, ncol = 12, byrow = TRUE)) %>% 
      setNames(1:12) %>%
    mutate(year = as.character(1959:1997))
head(co2_wide)
```

Use the gather() function to make this dataset tidy. Call the column with the CO2 measurements co2 and call the month column month. Name the resulting object co2_tidy.

Which code would return the correct tidy format?
```{r}
co2_tidy <- co2_wide %>% gather(key = "month",value = "co2","1":"12")
co2_tidy
```

Question 11
1 point possible (graded)
Use co2_tidy to plot CO2 versus month with a different curve for each year:

```{r}
co2_tidy %>% ggplot(aes(as.numeric(month), co2, color = year)) + geom_line()

```

What can be concluded from this plot? <br>
CO2 concentrations are highest around May and the yearly average increased from 1959 to 1997.<br>

---

### Question 12
Load the admissions dataset from dslabs, which contains college admission information for men and women across six majors, and remove the applicants percentage column:

```{r}
library(dslabs)
data(admissions)
dat <- admissions %>% select(-applicants)
head(dat)
```

Your goal is to get the data in the shape that has one row for each major, like this:
```{}
major  men   women
A      62    82		
B      63    68		
C      37    34		
D      33    35		
E      28    24		
F       6     7	
```
Which command could help you to wrangle the data into the desired format?
```{r}
dat %>% spread (gender, value = admitted)
```

---

### Question 13
Now use the admissions dataset to create the object tmp, which has columns major, gender, key and value:

```{r}
tmp <- gather(admissions, key, value, admitted:applicants)
tmp
```
    

Combine the key and gender and create a new column called column_name to get a variable with the following values: admitted_men, admitted_women, applicants_men and applicants_women. Save the new data as tmp2.

Which command could help you to wrangle the data into the desired format?
```{r}
tmp2 <- tmp %>% unite(column_name, key, gender)
tmp2
```

---

### Question 14
1 point possible (graded)
Which function can reshape tmp2 to a table with six rows and five columns named major, admitted_men, admitted_women, applicants_men and applicants_women?

```{}
major   admitted_men    admitted_women    applicants_men    applicants_women
A                              
B
```

```{r}
tmp2 %>% spread(key = column_name, value = value)
```


