---
title: "Exercises 02"
output: html_notebook
---

Question 1
1 point possible (graded)
You have the following table, schedule:

```{r}
schedule <- data.frame(c("Monday", "Tuesday"), c("Mandy, Chris and Laura","Steve, Ruth and Frank" ))
names(schedule) <- c("day", "staff")
schedule
```
You want to turn this into a more useful data frame
Which two commands would properly split the text in the “staff” column into each individual name? Select ALL that apply.

```{r}
str_split(schedule$staff, ", | and ")
```
```{r}
str_split(schedule$staff, ",\\s|\\sand\\s")
```

Question 2
1 point possible (graded)
You have the following table, schedule:

```{r}
schedule
```

What code would successfully turn your “Schedule” table into the following tidy table?
tidy

```
day     staff
<chr>   <chr>
Monday  Mandy
Monday  Chris
Monday  Laura
Tuesday Steve
Tuesday Ruth 
Tuesday Frank
```

```{r}
tidy <- schedule %>%  mutate(staff = str_split(staff, ", | and ")) %>%  unnest()
tidy
```


## Assessment Part 2: String Processing Part 3
Import raw Brexit referendum polling data from Wikipedia:

```{r, include =  FALSE}
library(rvest)
library(tidyverse)
library(stringr)
```

```{r}
url <- "https://en.wikipedia.org/w/index.php?title=Opinion_polling_for_the_United_Kingdom_European_Union_membership_referendum&oldid=896735054"
tab <- read_html(url) %>% html_nodes("table")
polls <- tab[[6]] %>% html_table(fill = TRUE)
polls
```


You will use a variety of string processing techniques learned in this section to reformat these data.

## Question 4
1 point possible (graded)
Some rows in this table do not contain polls. You can identify these by the lack of the percent sign (%) in the Remain column.

Update polls by changing the column names to c("dates", "remain", "leave", "undecided", "lead", "samplesize", "pollster", "poll_type", "notes") 

How many rows remain in the polls data frame?
```{r}
names(polls) <- c("dates", "remain", "leave", "undecided", "lead", "samplesize", "pollster", "poll_type", "notes")
polls
```
and only keeping rows that have a percent sign (%) in the remain column.
```{r}
d <- "%"
x <- filter(polls,  str_detect(remain, d) == TRUE)
x
```



```{r}
as.numeric(str_replace(x$remain, "%", ""))/100
```
```{r}
parse_number(x$remain)/100
```

## Question 6
3 points possible (graded)
The undecided column has some "N/A" values. These "N/A"s are only present when the remain and leave columns total 100%, so they should actually be zeros.

Use a function from stringr to convert "N/A" in the undecided column to 0. The format of your command should be function_name(polls$undecided, "arg1", "arg2").

What function replaces function_name?
```{r}
#polls$Undecided
str_replace(polls$Undecided, "N/A", "0")
```


## Question 7
0.0/3.5 points (graded)
The dates column contains the range of dates over which the poll was conducted. The format is "8-10 Jan" where the poll had a start date of 2016-01-08 and end date of 2016-01-10. Some polls go across month boundaries (16 May-12 June).

The end date of the poll will always be one or two digits, followed by a space, followed by the month as one or more letters (either capital or lowercase). In these data, all month abbreviations or names have 3, 4 or 5 letters.

Write a regular expression to extract the end day and month from dates. Insert it into the skeleton code below:
```{r}
(polls$`Date(s) conducted`)
```



```{r}
temp1 <- str_extract_all(polls$`Date(s) conducted`, "\\d+\\s[a-zA-Z]+")
temp2 <- str_extract_all(polls$`Date(s) conducted`, "\\d{1,2}\\s[a-zA-Z]+")
temp3 <- str_extract_all(polls$`Date(s) conducted`, "\\d+\\s[a-zA-Z]{3,5}")

identical(temp1, temp2)
identical(temp1, temp3)
identical(temp2, temp3)

#end_date <- sapply(temp, function(x) x[length(x)]) # take last element (handles polls that cross month boundaries)
```
      

Which of the following regular expressions correctly extracts the end day and month when inserted into the blank in the code above?

