---
title: "R Notebook"
output: html_notebook
---
Question 4
1 point possible (graded)
You have a data frame of monthly sales and profits in R:
```{}
Month     Sales     Profit 
<chr>     <chr>     <chr>  
January   $128,568  $16,234
February  $109,523  $12,876
March     $115,468  $17,920
April     $122,274  $15,825
May       $117,921  $15,437
```

Which of the following commands could convert the sales and profits columns to numeric? Select all that apply.

```{r}
#create df
dat <- data.frame(Month = c("Jan", "Feb", "Mar", "Apr", "May"), Sales = c("$128,568", "$109,523","$115,468","$122,274","$117,921"), Profit = c("$16,234","$12,876", "$17,920","$15,825","$15,437"))
dat
```
```{r}
a <- str_replace_all(dat$Sales, "\\$", "") 
a
```



convert the sales and profits columns to numeric
```{r}

dat %>% mutate_at(2:3, parse_number)
```
```{r}
dat %>% mutate_at(2:3, funs(str_replace_all(., c("\\$|,"), ""))) %>%   mutate_at(2:3, as.numeric)
```
### Question 12
1 point possible (graded)
In our example, we use the following code to detect height entries that do not match our pattern of x’y”:

```{r, include=FALSE}
library (tidyverse)
library (dslabs)
library (rvest)
library (htmlwidgets)
```

```{r}
data("reported_heights")
```


```{r}
not_inches <- function(x, smallest = 50, tallest = 84) {
  inches <- suppressWarnings(as.numeric(x))
  ind <- is.na(inches) | inches < smallest | inches > tallest 
  ind
}
```

```{r}
# identify entries with problems
problems <- reported_heights %>% 
  filter(not_inches(height)) %>%
  .$height
length(problems)
```



```{r}
converted <- problems %>% 
  str_replace("feet|foot|ft", "'") %>% 
  str_replace("inches|in|''|\"", "") %>% 
  str_replace("^([4-7])\\s*[,\\.\\s+]\\s*(\\d*)$", "\\1'\\2")



pattern <- "^[4-7]\\s*'\\s*\\d{1,2}$"
index <- str_detect(converted, pattern)
converted[index]
#converted[!index]
```

Which answer best describes the differences between the regex string we use as an argument in

```{r}
str_replace("^([4-7])\\s*[,\\.\\s+]\\s*(\\d*)$", "\\1'\\2") and the regex string in pattern <- "^[4-7]\\s*'\\s*\\d{1,2}$"?
```

### Question 13
1 point possible (graded)
You notice a few entries that are not being properly converted using your str_replace() and str_detect() code:
```{r}
yes <- c("5 feet 7inches", "5 7")
no <- c("5ft 9 inches", "5 ft 9 inches")
s <- c(yes, no)

converted <- s %>% 
  str_replace("\\s*(feet|foot|ft)\\s*", "'") %>% 
  str_replace("\\s*(inches|in|''|\")\\s*", "") %>% 
  str_replace("^([4-7])\\s*[,\\.\\s+]\\s*(\\d*)$", "\\1'\\2")

converted

#pattern <- "^[4-7]\\s*'\\s*\\d{1,2}$"
#str_detect(converted, pattern)
```

      
## Question 1
1 point possible (graded)
You have the following table, schedule:

```{r}
>schedule
day		staff
Monday		Mandy, Chris and Laura
Tuesday		Steve, Ruth and Frank
```


You want to turn this into a more useful data frame.

Which two commands would properly split the text in the “staff” column into each individual name? Select ALL that apply.

```{r}

```





    



